\section{Neptune} \label{neptune}

Neptune is our parallel answer to Julia's single-threaded garbage collector.
We use a fixed number of threads for marking and a single thread for sweeping, for reasons to be explained shortly. 
Its overall structure is highly similar to Julia's for several reasons (as can be seen in figure \ref{fig:stages:np}).
First, our goal was to test the effects of parallelism on garbage collection, not to experiment with novel approaches to the general garbage collection problem.
Second, Julia's C implementation is highly complex, relying on delicate memory layout requirements, among other things.
We did not want to make our job of understanding, tracking, marking, and sweeping the contents of memory as organized originally by Julia any harder by attempting optimizations on their algorithm, apart from the complexities of adding parallelism.

\subsection{Why Rust}
Rust is a modern systems programming language, created in 2010.
It prides itself in providing many ``zero-cost'' abstractions -- abstractions and guarantees that in other systems language normally require extensive library support or checks.
It gives the user memory safety and automatic memory management \textbf{without} garbage collection due to its sophisticated type system, which incorporates ``affine'' types.
All of this makes concurrency quick and easy in Rust, and also importantly, it provides a nice foreign function interface, which would prove essential for interacting with Julia.

\subsection{Design Decisions}
Once we completed our implementation of Julia's garbage collector in Rust (which took the bulk of our time), we had to decide where to proceed in achieving our goal of a parallel garbage collector.
Our general design was inspired by \cite{marlow2008parallel} and \cite{marlow2011multicore}, especially the parts about thread-local data structures and handling parallelism.
Our specific design decisions were primarily profile-guided, meaning we used the output of tools like Valgrind and OProfile, as well as in-code timers and measurements, to determine where to prioritize parallelization implementation.
We saw that while marking would benefit greatly from having multiple threads, parallel sweeping actual hindered performance.
Since sweeping is mostly memory-bound, parallelizing it increased cache misses.
We also took advantage of Rust's libaries in heavily using lock-free data structures and caches in many places to prevent unnecessary blocking and lock contention.

Neptune reads the value of the environment variable \texttt{NEPTUNE\_THREADS} and creates a work-stealing thread pool.
The parallel marking algorithm described in section~\ref{marking_algo} uses this thread pool rather than creating new threads every single time to avoid the overhead of thread creation.

\subsection{The Parallel Marking Algorithm}
\label{marking_algo}
We adapted Julia's marking algorithm to be multithreaded.
The original marking algorithm is implemented as a recursive algorithm that walks the root set and pushes all the leaves onto a stack (called the \emph{mark stack}) when it reaches a certain recursion depth.
To parallelize this algorithm, we made a couple changes:
Firstly, we replaced the original mark stack with a Treiber stack, a simple lock-free, thread-safe, and memory-safe stack algorithm.
Secondly, we enhanced all marking with atomic updates to be thread-safe.
Finally, we implemented some thread-local caches that get synchronized after marking to prevent having too many expensive accesses to global data structures accompanied with locks.
To do so, we extended Julia's mark caches and added one thread-local mark cache per garbage collection worker thread.
Our thread-local mark caches contain local updates to remsets, big object lists, and statistics for future collection decisions.

Our marking algorithm gives marking jobs for all objects in the mark stack to the thread pool and waits for the worker threads to finish and synchronize.
If the mark stack is not empty because worker threads added more objects in the meantime, the algorithm will do more iterations of the same job assignment until the mark stack is empty after synchronization.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:
